\documentclass{ximera}
\input{../preamble}
\title{Exercises: General Slicing}
\author{Philip T. Gressman}

\begin{document}
\begin{abstract}
Exercises computing volume by cross-sectional area.
\end{abstract}
\maketitle

\begin{exercise}
The base of a solid region is bounded by the curves $x = 0$, $y = x^2$ and $y = x$. The cross sections perpendicular to the $x$-axis are squares. Compute the volume of the region.
\begin{itemize}
\item A typical square cross section has side length $L = \answer{x - x^2}$ and area $A = \answer{(x-x^2)^2}$. 
\item Possible numerical values of the $x$-coordinates of points in the base range from a minimum value of $x = \answer{0}$ up to a maximum of $x = \answer{1}$.
\item To compute volume, integrate:
\[ V = \int_{\answer{0}}^{\answer{1}} \answer{(x-x^2)^2} d\answer{x} = \answer{\frac{1}{30}}. \]
\end{itemize}
\end{exercise}

\begin{exercise}
Find the volume of the region in three-dimensional space defined by the inequalities
\begin{align*}
0 & \leq x \leq 1, \\
0 & \leq y \leq z^2, \\
0 & \leq z \leq 3.
\end{align*}
\begin{itemize}
\item Cross sections perpendicular to the $z$-axis are \wordChoice{\choice{square}\choice[correct]{rectangular}\choice{triangluar}} with length $\answer{1}$ in the $x$-direction and width $\answer{z^2}$ in the $y$-direction.
\item The area of a $z$ cross section is $A(z) = \answer{z^2}$.
\item To compute volume, integrate:
\[ V = \int_{\answer{0}}^{\answer{3}} \answer{z^2} dz = \answer{9}. \]
\end{itemize}
\end{exercise}

\begin{exercise}
A right circular cylinder of radius $1$ and height $3$ is twisted along its axis so that the disk at height $z$ is centered on the axis $ x = \cos (2 \pi z/3), y = \sin (2 \pi z/3))$, which corresponds to one full twist along the axis. Compute the volume of this twisted cylinder.
\[ V = \answer{3 \pi}. \]
\end{exercise}

\begin{exercise}
A certain three-dimensional region has a base in the $xy$-plane which is bounded above by the graph $y = 1-x^2$ and below by $y=0$. Slices perpendicular to the $y$-axis are equilateral triangles whose base lies in the $xy$-plane as well. Compute the volume of the region.
\[ V = \answer{\frac{\sqrt{3}}{2}}. \]
\end{exercise}

\section*{Sample Exam Questions}

\begin{question}(2018 Midterm 1)
Compute the volume of the region in 3-dimensional space which satisfies the inequalities
\begin{align*} 
0 \leq   x \leq (1 - z^2) \qquad \mbox{ and } \qquad
0 \leq  y \leq (1 + z^2) \qquad \mbox{ and } \qquad 
0 \leq z \leq 1.
\end{align*}
\vspace{-12pt}
\begin{multipleChoice}
\choice{$\displaystyle \frac{2}{3}$}
\choice{$\displaystyle \frac{3}{4}$}
\choice[correct]{$\displaystyle \frac{4}{5}$}
\choice{$\displaystyle \frac{5}{6}$}
\choice{$\displaystyle \frac{6}{7}$}
\choice{none of these}
\end{multipleChoice}
\end{question}

\begin{question}(2019 Midterm 1)
The inequality 
\[ \frac{x^2}{a^2} + \frac{y^2}{b^2} \leq 1 \]
defines an ellipse in the $xy$-plane whose area is $\pi a b$ for any positive values of the constants $a$ and $b$.  Compute the three dimensional volume of the region defined by
\[ 4 x^2 + z^2 y^2 \leq z^2 \mbox{ for } 0 \leq z \leq 1. \]
\begin{multipleChoice}
\choice{$\displaystyle 4 \pi z^2$}
\choice{$4 \pi$}
\choice{$\displaystyle \frac{\pi}{4z}$}
\choice[correct]{$\displaystyle \frac{\pi}{4}$}
\choice{$\displaystyle {\pi z}$}
\choice{$\displaystyle {\pi}$}
\end{multipleChoice}
Comments and hints for after your first answer:
\begin{feedback}
Dividing the first inequality by $z^2$ on both sides gives
\[ \frac{x^2}{\frac{z^2}{4}} + \frac{y^2}{1} \leq 1, \]
\begin{hint}
which means that slices in the $z$-direction are ellipses with area $\pi \frac{z}{2} \cdot 1 = \frac{\pi z}{2}$. 
\end{feedback}
\begin{feedback}
Volume is obtained by integration:
\[ V = \int_0^1 \frac{\pi z}{2} dz = \left. \frac{\pi z^2}{4} \right|_{0}^1 = \frac{\pi}{4}. \]
\end{hint}
\end{feedback}
\end{question}


\end{document}
