\documentclass{ximera}
\input{../preamble}
\title{Separable and Linear ODEs}
%%%%%\author{Philip T. Gressman}

\begin{document}
\begin{abstract}
We learn techniques to solve first-order linear and separable ODEs.
\end{abstract}
\maketitle

\section*{Online Texts}
\begin{itemize}
\item \link[OpenStax II 4.3: Separable Equations]{https://openstax.org/books/calculus-volume-2/pages/4-3-separable-equations} and \link[Linear Equations]{https://openstax.org/books/calculus-volume-2/pages/4-5-first-order-linear-equations}
\item \link[Ximera OSU: Separable Equations]{https://ximera.osu.edu/mooculus/calculus2/separableDifferentialEquations/titlePage}
\end{itemize}

\section*{Examples}

\begin{example}
The ODE
\[ y' =  x^2 e^y \]
is separable. 
\begin{itemize}
\item To separate, we move all factors involving $y$ to the left-hand side and all factors involving $x$ to the right-hand side. We also formally write $y'$ as $dy/dx$ and verify that we can end up with $dy$ on the left-hand side and $dx$ on the right hand side:
\[ \answer{e^{-y}} dy = \answer{x^2} dx. \]
\item Apply an integral on both sides:
\[ \int  \answer{e^{-y}} dy = \int \answer{x^2} dx \Rightarrow \answer{-e^{-y}} = \answer{\frac{x^3}{3}} + C.\]
That gives our general solution.
\item If we had initial conditions $y(0) = 1$, we would plug in $x=\answer{0}$, $y=\answer{1}$ to our general solution to determine that $C = \answer{-e^{-1}}$.  We can plug in this value for $C$ and solve for $y$ to determine that
\[ y = \answer{- \ln \left( e^{-1} - \frac{x^3}{3} \right)}. \]
\end{itemize}
\end{example}

\begin{example}
The ODE
\[ x y' = -x^2 y +  e^{-x^2/2}\]
is linear.
\begin{itemize}
\item To begin, we must write the ODE in standard form:
\[ y' + \answer{x} y = \answer{e^{-x^2/2}/x}. \]
\item We give the name $P(x)$ to the coefficient of $y$ (note that it doesn't \textit{have to} depend on $x$. Writing $P(x)$ is only meant to indicate that it \textit{cannot} depend on $y$). We give the name $Q(x)$ to the right-hand side.
\item The integrating factor is the exponential of the integral of $P$. In this case, we can take $I(x) = e^{x^2/2}$. Multiply both sides of the standard form equation by the integrating factor:
\[ \answer{e^{x^2/2}} y' + \answer{x e^{x^2/2}} y = \answer{\frac{1}{x}}. \]
\item The integrating factor gets its name because it is now possible to integrate the left-hand side.  By the product rule, we know that
\[ \left( e^{x^2/2} y \right)' = \answer{e^{x^2/2}} y' + \answer{x e^{x^2/2}} y, \]
which is exactly the right-hand side. Thus we have
\[ \left( e^{x^2/2} y \right)' = \answer{\frac{1}{x}}. \]
\item Integrating the left-hand side gives $e^{x^2/2} y$; integrating both sides gives
\[ e^{x^2/2} y = \answer{\ln |x|} + C \]
(don't forget absolute values in the logarithm).  Solving for $y$ gives
\[ y = \answer{e^{-x^2/2} \ln |x|} + C e^{-x^2/2} \]
as the general solution.
\end{itemize}
\end{example}


\end{document}
