\documentclass{ximera}
\input{../preamble}
\title{Integration by Parts}
\author{Philip T. Gressman}

\begin{document}
\begin{abstract}
We study the integration technique of integration by parts.
\end{abstract}
\maketitle

\begin{example}
Compute the indefinite integral
\[ \int x^4 e^{2x} ex. \]
\begin{itemize}
\item We'll do this using an organizational technique called ``tabular integration'' that many people find helpful when doing repeated integrations by parts.
\item Make a table: one column for $u$ and another for $dv$. In the first row, rewrite the functions that you will use for $u$ and $dv$:
\begin{center}
\begin{tabular}{ccc}
 & $u$ & $dv$ \\
 \hline
& $\displaystyle x^4$ & $\displaystyle e^{2x}$ 
\end{tabular}
\end{center}
\item Add new rows: differentiate items in the $u$ column and integrate items in the $dv$ column to determine what items go in the next row. 
\begin{center}
\begin{tabular}{ccc}
 & $u$ & $dv$ \\
 \hline
& $\displaystyle x^4$ & $\displaystyle e^{2x}$ \\
& $\displaystyle 4 x^3$ & $\displaystyle e^{2x}/2$ \\
& $\displaystyle \answer{12 x^2}$ & $\displaystyle \answer{e^{2x}/4}$ \\
& $\displaystyle \answer{24 x}$ & $\displaystyle \answer{e^{2x}/8}$ \\
& $\displaystyle \answer{24}$ & $\displaystyle  \answer{e^{2x}/16}$ \\
\end{tabular}
\item Now you combine terms by matching items in the first column with items \textit{one row down}. For the last item in the $u$ column, match it with the last item in the $dv$ column (which you'll end up using twice):
\[ x^4 \cdot \frac{e^{2x}}{2} \qquad {4 x^3} \cdot \answer{\frac{e^{2x}}{4}} \qquad \answer{12 x^2} \cdot \answer{\frac{e^{2x}}{8}} \qquad \answer{24 x} \cdot \answer{\frac{e^{2x}}{16}} \qquad \answer{24} \cdot \answer{\frac{e^{2x}}{16}} \]
\end{center}
\item To finish, you alternate addition and subtraction. Give a $+$ to the first term, a $-$ to the second, and so on. Last but not least, put an integral on the last term as well:
\[ + x^4 \frac{e^{2x}}{2} - \answer{4 x^3 \frac{e^{2x}}{4}} + \answer{12 x^2 \frac{e^{2x}}{8}} - \answer{24 x\frac{e^{2x}}{16} }+ \int \answer{24 \frac{e^{2x}}{16}} dx \]
In this way, we arrive at the formula
\[ \int x^4 e^{2x} dx = \frac{1}{2} x^4 e^{2x} - x^3 e^{2x} + \frac{3}{2} x^2 e^{2x} - \frac{3}{2} x e^{2x} + \int \frac{3}{2} e^{2x} dx \]
\item Sometimes the integrand of the last term is zero. This would happen when $u$ is a polynomial if you have enough lines in your table. In that case there would be nothing else to compute.  (We stopped one line short of that point on purpose in this example to show how you would handle a case where there the last term \textit{isn't} zero.)
\end{itemize}
\end{example}
\begin{example}

\end{example}


\end{document}
